---
title: "Data Exploration"
author: "Lidio Jaimes Jr"
date: "2024-10-21"
output: pdf_document
---
```{r Table 1: Patient Demographics}
label(stroke$race) <- "Race"
label(stroke$age) <- "Age"
label(stroke$sex) <- "Sex"
label(stroke$bmi_kg_m_2) <- "BMI"
label(stroke$cholesterol_mg_d_l) <- "Cholesterol"
label(stroke$admission_nihss) <- "Admission NIHSS"
label(stroke$prior_stroke) <- "Prior Stroke"
label(stroke$length_of_admission_days) <- "Length of Admission"
label(stroke$covid_19) <- "COVID-19"
label(stroke$caregiver) <- "Presence of Caregiver"


units(stroke$age) <- "years"
units(stroke$bmi_kg_m_2) <- "kg/m^2"
units(stroke$cholesterol_mg_d_l) <- "mg/dL"
units(stroke$length_of_admission_days) <- "days"

table1(~ race + age + sex + bmi_kg_m_2 + cholesterol_mg_d_l + covid_19 +
         prior_stroke + caregiver + admission_nihss + length_of_admission_days | location_of_stroke,
       overall = c(left = "Overall"),
       data = stroke)
```

```{r Figure 1: Missingness}
stroke_variables<-
  stroke |> 
  select(race,age,sex,bmi_kg_m_2,cholesterol_mg_d_l,prior_stroke,admission_nihss,covid_19)

VIM::aggr(stroke_variables,
          col = c("blue","red"),
          numbers = TRUE,
          labels = c("Race","Age","Sex","BMI","Cholesterol",
                     "Prior Stroke","Entry NIHSS","COVID-19"),
          ylab = c("Histogram of missing data", "Pattern"))
```


```{r Figure 2 : Stroke location vs NIHSS}
stroke |> 
  ggplot(aes(x=admission_nihss,y=fct_rev(location_of_stroke),fill=location_of_stroke)) + 
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Admission NIHSS") + 
  ylab("Location of Stroke")
```

Using cutoffs from JHU, hopkinsmedicine.org
```{r Figure 3: Obesity}
stroke |> 
  ggplot(aes(x = admission_nihss, y = bmi_kg_m_2, color = cholesterol_group)) +
  scale_color_manual(name = "Cholesterol Group",
                     labels = c("Normal","Boderline High","High"),
                     values = c("red","blue", "green")) + 
  geom_point() + 
  labs(x = "Admission NIHSS", y = "BMI") + 
  theme_bw() +
  geom_hline(yintercept=18.5, linetype="dashed", color = "black") +
  geom_hline(yintercept=25, linetype="dashed", color = "black")+
  theme(legend.position = "bottom")

```

```{r Figure 4: Diabetes}
stroke |> 
  group_by(location_of_stroke,dm) |> 
  summarise(median_nihss=median(admission_nihss,na.rm=TRUE)) |> 
  filter(!is.na(dm)) |> 
  ggplot(aes(x = (location_of_stroke),
             y = median_nihss,
             fill = dm)) + 
  geom_bar(stat = "identity", position = "dodge", width = 0.7, color="black") +
  scale_fill_manual(name = "Diabetic",
                     labels = c("No","Yes"),
                     values = c("red","blue")) + 
  theme_bw() +
  xlab("Location of Stroke") + 
  ylab("Median Admission NIHSS Score")
```

```{r Figure 5: location of stroke vs age}
stroke |> 
  ggplot(aes(x=age,y=fct_rev(location_of_stroke),fill=location_of_stroke)) + 
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Age") + 
  ylab("Location of Stroke")
```

```{r Figure 6: Smoking history vs NIHSS}
stroke |> 
  ggplot(aes(x=admission_nihss,y=fct_rev(cigarette_smoking),fill=cigarette_smoking)) + 
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Admission NIHSS") + 
  ylab("Smoking History")
```

This is about COVID-19 variable: 
```{r Table 2: COVID-19}
stroke[stroke$id_number==48,"covid_19_notes"] = "2022-09-18"
stroke[stroke$id_number==139,"covid_19_notes"] = "2021-09-19"
stroke[stroke$id_number==143,"covid_19_notes"] = "2021-07-27"
stroke[stroke$id_number==168,"covid_19_notes"] = "2022-01-31"

stroke |> 
  mutate(covid_19=factor(covid_19,levels=c("Yes","No"))) |> 
  rename(`COVID-19`=covid_19) |> 
  group_by(`COVID-19`) |> 
  summarise(n())

#kableExtra::kable(ppt_tbl)
```

Patients 4, 48, 103, 139, 143, and 168 all have "Yes (Once)" for COVID-19 variable, even though they had a stroke prior to the COVID-19 date 2020-03-15. 

```{r Figure 7: COVID}
stroke |> 
  mutate(covid_19=case_when(covid_19=="Yes"~"Yes",covid_19=="No"~"No",is.na(covid_19)~"NA"),
         covid_19=factor(covid_19)) |> 
  ggplot(aes(x=admission_nihss,y=covid_19,fill=covid_19)) + 
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Admission NIHSS") + 
  ylab("COVID-19")
```

Is the COVID variable whether they have any history or if it's present when they had stroke? maybe a mix of both? Patient 72 tested positive for COVID on 7/21/21 and their date of stroke is 11/5/20. The COVID-19 notes is that the patient expired on 7/27/21 secondary to COVID-19. The cutoff date for COVID is 3/15/20

```{r Figure 8: Heart disease vs NIHSS}
stroke |> 
  ggplot(aes(x=admission_nihss,y=fct_rev(ischemic_heart_disease),fill=ischemic_heart_disease)) + 
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Admission NIHSS") + 
  ylab("Ischemic Heart Disease History")
```

```{r Figure 9: Depression vs Length of Admission}
stroke |> select(id_number,length_of_admission_days,date_of_stroke,expiration_date_in_hospital,expiration_date)

stroke |>
  mutate(expiration = factor(expiration, levels = c(0, 1), labels = c("No", "Yes"))) |> 
  filter(expiration=="No") |> 
  ggplot(aes(x=length_of_admission_days,y=depression,fill=depression))+
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Length of Admission (days)") + 
  ylab("Depression")

stroke |>
  mutate(expiration = factor(expiration, levels = c(0, 1), labels = c("No", "Yes"))) |> 
  filter(expiration=="No") |> 
  ggplot(aes(x=length_of_admission_days,y=depression,fill=depression))+
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Length of Admission (days)") + 
  ylab("Depression")


stroke |>
  mutate(expiration = factor(expiration, levels = c(0, 1), labels = c("No", "Yes"))) |> 
  filter(length_of_admission_days<75) |> 
  ggplot(aes(x=length_of_admission_days,y=depression,fill=expiration))+
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,alpha=.2)+
  theme_bw() +
  xlab("Length of Admission (days)") + 
  ylab("Depression")

stroke |>
  filter(length_of_admission_days<75) |> 
  ggplot()+
  geom_violin(aes(x=length_of_admission_days,y=depression,fill=expiration)) + 
  geom_boxplot(aes(x=length_of_admission_days,y=depression,fill=expiration),alpha=.2)+
  theme_bw() +
  xlab("Length of Admission (days)") + 
  ylab("Depression")
```

Patient 80 has length of admission days as 3, but they died 5 days after stroke. Patient 101. **NEED** to ask about length_of_admission_days variable. 

```{r Figure 10: Depression vs NIHSS}
stroke |> 
  ggplot(aes(x=admission_nihss,y=depression,fill=depression))+
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Admission NIHSS") + 
  ylab("Depression")
```

```{r Score completeness}
table(is.na(stroke$pt_am_pac_score))
table(is.na(stroke$admission_mrs_score))
table(is.na(stroke$admission_nihss))

stroke |> 
  filter(is.na(pt_am_pac_score),is.na(admission_mrs_score),is.na(admission_nihss)) |>
  select(id_number,date_of_stroke)
```

```{r Score relationship}
stroke |> 
  select(admission_nihss,admission_mrs_score,pt_am_pac_score,
         discharge_nihss,discharge_mrs_score,ot_am_pac_score) |> 
  rename(admit_nihss = admission_nihss, 
         admit_mrs = admission_mrs_score,
         pt_ampac = pt_am_pac_score,
         dis_nihss = discharge_nihss,
         dis_mrs = discharge_mrs_score,
         ot_ampac = ot_am_pac_score) |> 
  #na.omit() |> # 115 patients
  GGally::ggpairs()

stroke |> 
  ggplot(aes(x=admission_nihss,y=pt_am_pac_score)) + 
  geom_point()
```

```{r Score Correlation}
stroke |> 
  select(admission_nihss,admission_mrs_score,pt_am_pac_score,
         discharge_nihss,discharge_mrs_score,ot_am_pac_score) |> 
  rename(admit_nihss = admission_nihss, 
         admit_mrs = admission_mrs_score,
         pt_ampac = pt_am_pac_score,
         dis_nihss = discharge_nihss,
         dis_mrs = discharge_mrs_score,
         ot_ampac = ot_am_pac_score) |> 
  na.omit() |> # 115 patients
  cor() |> 
  corrplot::corrplot.mixed(tl.cex=.75)

stroke |> 
  select(admission_nihss,admission_mrs_score,pt_am_pac_score) |> 
  rename(admit_nihss = admission_nihss, 
         admit_mrs = admission_mrs_score,
         pt_ampac = pt_am_pac_score) |> 
  na.omit() |> # 266 patients
  cor() |> 
  corrplot::corrplot.mixed()

stroke |> 
  select(discharge_nihss,discharge_mrs_score,ot_am_pac_score) |> 
  rename(dis_nihss = discharge_nihss,
         dis_mrs = discharge_mrs_score,
         ot_ampac = ot_am_pac_score) |> 
  na.omit() |> # 123 patients
  cor() |> 
  corrplot::corrplot.mixed()
```

