---
title: "Data Exploration"
author: "Lidio Jaimes Jr"
date: "2024-10-21"
output: pdf_document
---
```{r Importing libraries}
library(tidyverse)
library(forcats)
library(table1)
```

```{r Reading in Data}
stroke_raw <- 
  readxl::read_xlsx("Capstone Combined Data Collection Sheets - Summer 2024.xlsx", 
                    skip = 1,
                    na=c("", " ", "UNK", "Unk", "NA", "na", "Na", 9999, "????")) |> 
  janitor::clean_names()
stroke <- stroke_raw[-c(1,376), ] #get rid of dictionary and black row

# E336 (patient 333) and E340 (patient 337) have full date instead of just year
stroke[stroke$id_number == 333, "year_of_birth"] = 1905 #set as 1981 instead of 1905
stroke[stroke$id_number == 337, "year_of_birth"] = 1942
stroke$date_identified = openxlsx::convertToDate(stroke$date_identified)
stroke$date_of_stroke = openxlsx::convertToDate(stroke$date_of_stroke)
stroke$expiration_date_in_hospital = openxlsx::convertToDate(stroke$expiration_date_in_hospital)
stroke$expiration_date = openxlsx::convertToDate(stroke$expiration_date)



#Patients 211, 391 (repeat MRN), 528, 529, 536, 537, 555, 558, 576, 582
stroke <- 
  stroke |> 
  filter(!id_number%in%c(211, 391, 528, 529, 536, 537, 555, 558, 576, 582),
         ischemic == 1) #focus on ischemic
#patient 100 has a good bit of missing
```

```{r Data Wrangling}
stroke$brainstem_stroke_lcation <- as.numeric(stroke$brainstem_stroke_lcation)

stroke <-
  stroke |>
  mutate(multiple_locations = ifelse(rowSums(stroke[,106:110], na.rm = TRUE) > 1, 1, 0),
         location_of_stroke = 
           case_when(
             multiple_locations==0 & brainstem_stroke_lcation == 1 ~ "brainstem",
             multiple_locations==0 & cerebellum_stroke_location == 1 ~ "cerebellum",
             multiple_locations==0 & cortical_stroke_location == 1 ~ "cortical",
             multiple_locations==0 & subcortical_stroke_location == 1 ~ "subcortical",
             multiple_locations==0 & other_stroke_location == 1 ~ "other/unknown",
             multiple_locations==1 ~ "multiple locations",
             TRUE ~ "other/unknown"),
         location_of_stroke = factor(location_of_stroke,
                                     levels = c("brainstem","cerebellum",
                                                "cortical","subcortical",
                                                "other/unknown","multiple locations"),
                                     labels = c("Brainstem","Cerebellum",
                                                "Cortical","Subcortical",
                                                "Unknown","Multiple Locations")),
         race = ifelse(race=="Other","Other","White"),
         race = case_when(
           race == "White" ~ "White",
           str_detect(patient_identification_notes, "African") ~ "Black",
           TRUE ~ "Other"),
         race = factor(race, levels = c("White", "Black", "Other")),
         age = as.numeric(age),
         admission_nihss = as.numeric(admission_nihss),
         cholesterol_mg_d_l = as.numeric(cholesterol_mg_d_l),
         prior_stroke = case_when(
           prior_stroke == 0 ~ "No",
           prior_stroke == 1 ~ "Yes"),
         prior_tia = case_when(
           prior_tia == 0 ~ "No",
           prior_tia == 1 ~ "Yes"),
         cholesterol_group = case_when(
           cholesterol_mg_d_l < 200 ~ "Normal",
           cholesterol_mg_d_l >= 200 & cholesterol_mg_d_l <=239 ~ "Borderline high",
           cholesterol_mg_d_l >= 240 ~ "High"),
         cholesterol_group = factor(cholesterol_group,
                                    levels = c("Normal","Borderline high","High")),
         cigarette_smoking = case_when(
           cigarette_smoking == "Former" | cigarette_smoking == "former" ~ "Former",
           cigarette_smoking == "Never" | cigarette_smoking == "NEver" ~ "Never",
           cigarette_smoking == "Current" | cigarette_smoking == "Current\r\n\rCurrent" ~ "Current"),
         cigarette_smoking = factor(cigarette_smoking,
                                    levels = c("Never","Former","Current")),
         ischemic_heart_disease = ifelse(ischemic_heart_disease == 0, "No", "Yes"),
         date_cov19 = ifelse(date_of_stroke>="2020-03-15","after covid","before covid"),
         covid_19 = ifelse(date_cov19=="before covid","No",covid_19),
         covid_19 = case_when(
           covid_19 == "No" ~ "No",
           covid_19 == "Yes (Once)" ~ "Yes",
           covid_19 == "positive" ~ "Yes",
           covid_19 == "negative" ~ "No"),
         dm = factor(dm,levels=c(0,1),labels=c("No","Yes")),
         depression = factor(depression,levels=c(0,1),labels=c("No","Yes")),
         length_of_admission_days = as.numeric(length_of_admission_days))

stroke[stroke$id_number==532,"race"] = "Other"
stroke[stroke$id_number==223,"race"] = "Other"
```

Patients with Race classified as other are either: Other(Native American and Native Hawaiian), refused to answer question, or are mixed (patient 532 and 223). No one has a stroke_location of other, people with missing data were grouped here. Some of the people with no stroke_location, have under the notes column that it is possible TIA, should these people be excluded from analysis? 

This is about COVID-19 variable: 
```{r COVID-19}
stroke[stroke$id_number==48,"covid_19_notes"] = "2022-09-18"
stroke[stroke$id_number==139,"covid_19_notes"] = "2021-09-19"
stroke[stroke$id_number==143,"covid_19_notes"] = "2021-07-27"
stroke[stroke$id_number==168,"covid_19_notes"] = "2022-01-31"

stroke |> 
  mutate(covid_19=factor(covid_19,levels=c("Yes","No"))) |> 
  rename(`COVID-19`=covid_19) |> 
  group_by(`COVID-19`) |> 
  summarise(n())

#kableExtra::kable(ppt_tbl)
```

Patients 4, 48, 103, 139, 143, and 168 all have "Yes (Once)" for COVID-19 variable, even though they had a stroke prior to the COVID-19 date 2020-03-15. 

```{r Table 1: Patient Demographics}
label(stroke$race) <- "Race"
label(stroke$age) <- "Age"
label(stroke$sex) <- "Sex"
label(stroke$bmi_kg_m_2) <- "BMI"
label(stroke$cholesterol_mg_d_l) <- "Cholesterol"
label(stroke$admission_nihss) <- "Admission NIHSS"
label(stroke$prior_stroke) <- "Prior Stroke"
units(stroke$age) <- "years"
units(stroke$bmi_kg_m_2) <- "kg/m^2"
units(stroke$cholesterol_mg_d_l) <- "mg/dL"

table1(~ race + age + sex + bmi_kg_m_2 + cholesterol_mg_d_l + 
         prior_stroke + admission_nihss | location_of_stroke,
       overall = c(left = "Overall"),
       data = stroke)
```

```{r Figure 1: Missingness}
stroke_variables<-
  stroke |> 
  select(race,age,sex,bmi_kg_m_2,cholesterol_mg_d_l,prior_stroke,admission_nihss,covid_19)

VIM::aggr(stroke_variables,
          col = c("blue","red"),
          numbers = TRUE,
          labels = c("Race","Age","Sex","BMI","Cholesterol",
                     "Prior Stroke","Entry NIHSS","COVID-19"),
          ylab = c("Histogram of missing data", "Pattern"))
```


```{r Figure 2 : Stroke location vs NIHSS}
stroke |> 
  ggplot(aes(x=admission_nihss,y=fct_rev(location_of_stroke),fill=location_of_stroke)) + 
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Admission NIHSS") + 
  ylab("Location of Stroke")
```

Using cutoffs from JHU, hopkinsmedicine.org
```{r Figure 3: Obesity}
stroke |> 
  ggplot(aes(x = admission_nihss, y = bmi_kg_m_2, color = cholesterol_group)) +
  scale_color_manual(name = "Cholesterol Group",
                     labels = c("Normal","Boderline High","High"),
                     values = c("red","blue", "green")) + 
  geom_point() + 
  labs(x = "Admission NIHSS", y = "BMI") + 
  theme_bw() +
  geom_hline(yintercept=18.5, linetype="dashed", color = "black") +
  geom_hline(yintercept=25, linetype="dashed", color = "black")+
  theme(legend.position = "bottom")

```

```{r Figure 4: Diabetes}
stroke |> 
  group_by(location_of_stroke,dm) |> 
  summarise(median_nihss=median(admission_nihss,na.rm=TRUE)) |> 
  filter(!is.na(dm)) |> 
  ggplot(aes(x = (location_of_stroke),
             y = median_nihss,
             fill = dm)) + 
  geom_bar(stat = "identity", position = "dodge", width = 0.7, color="black") +
  scale_fill_manual(name = "Diabetic",
                     labels = c("No","Yes"),
                     values = c("red","blue")) + 
  theme_bw() +
  xlab("Location of Stroke") + 
  ylab("Median Admission NIHSS Score")
```

```{r Figure 5: location of stroke vs age}
stroke |> 
  ggplot(aes(x=age,y=fct_rev(location_of_stroke),fill=location_of_stroke)) + 
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Age") + 
  ylab("Location of Stroke")
```

```{r Figure 6: Smoking history vs NIHSS}
stroke |> 
  ggplot(aes(x=admission_nihss,y=fct_rev(cigarette_smoking),fill=cigarette_smoking)) + 
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Admission NIHSS") + 
  ylab("Smoking History")
```

```{r Figure 7: COVID}
stroke |> 
  mutate(covid_19=case_when(covid_19=="Yes"~"Yes",covid_19=="No"~"No",is.na(covid_19)~"NA"),
         covid_19=factor(covid_19)) |> 
  ggplot(aes(x=admission_nihss,y=covid_19,fill=covid_19)) + 
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Admission NIHSS") + 
  ylab("COVID-19")
```

Is the COVID variable whether they have any history or if it's present when they had stroke? maybe a mix of both? Patient 72 tested positive for COVID on 7/21/21 and their date of stroke is 11/5/20. The COVID-19 notes is that the patient expired on 7/27/21 secondary to COVID-19. The cutoff date for COVID is 3/15/20

```{r Figure 8: Heart disease vs NIHSS}
stroke |> 
  ggplot(aes(x=admission_nihss,y=fct_rev(ischemic_heart_disease),fill=ischemic_heart_disease)) + 
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Admission NIHSS") + 
  ylab("Ischemic Heart Disease History")
```

```{r Figure 9: Depression vs Length of Admission}
stroke |> select(id_number,length_of_admission_days,date_of_stroke,expiration_date_in_hospital,expiration_date)

stroke |>
  mutate(expiration = factor(expiration, levels = c(0, 1), labels = c("No", "Yes"))) |> 
  filter(expiration=="No") |> 
  ggplot(aes(x=length_of_admission_days,y=depression,fill=depression))+
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Length of Admission (days)") + 
  ylab("Depression")

stroke |>
  mutate(expiration = factor(expiration, levels = c(0, 1), labels = c("No", "Yes"))) |> 
  filter(expiration=="No") |> 
  ggplot(aes(x=length_of_admission_days,y=depression,fill=depression))+
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Length of Admission (days)") + 
  ylab("Depression")


stroke |>
  mutate(expiration = factor(expiration, levels = c(0, 1), labels = c("No", "Yes"))) |> 
  filter(length_of_admission_days<75) |> 
  ggplot(aes(x=length_of_admission_days,y=depression,fill=expiration))+
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,alpha=.2)+
  theme_bw() +
  xlab("Length of Admission (days)") + 
  ylab("Depression")

stroke |>
  filter(length_of_admission_days<75) |> 
  ggplot()+
  geom_violin(aes(x=length_of_admission_days,y=depression,fill=expiration)) + 
  geom_boxplot(aes(x=length_of_admission_days,y=depression,fill=expiration),alpha=.2)+
  theme_bw() +
  xlab("Length of Admission (days)") + 
  ylab("Depression")
```

Patient 80 has length of admission days as 3, but they died 5 days after stroke. Patient 101. **NEED** to ask about length_of_admission_days variable. 

```{r Figure 10: Depression vs NIHSS}
stroke |> 
  ggplot(aes(x=admission_nihss,y=depression,fill=depression))+
  geom_violin(alpha=.5) + 
  geom_boxplot(width=.15,fill="grey",alpha=.2)+
  theme_bw() +
  theme(legend.position = "none") + 
  xlab("Admission NIHSS") + 
  ylab("Depression")
```



```{r Discharge} 
discharge_info <- 
  stroke |> 
  select(id_number,expiration,acute_rehabilitation,subacute_rehabilitation,
         inpatient_rehabilitation,assited_living_facility,home_health,
         home_with_no_needs,hospice,other_discharge_location) |> 
  sapply(as.numeric) |> 
  as.data.frame()

discharge_info |> 
  mutate(sum_discharge=apply(discharge_info[3:9],1,sum)) |> 
  filter(sum_discharge==2)


stroke[, c(172,174:177,179:182)] <- lapply(stroke[, c(172,174:177,179:182)], as.numeric)

test<-
  stroke |> 
  mutate(multiple_discharge = 
           ifelse(rowSums(stroke[,c(172,174:177,179:182)], na.rm = TRUE) > 1, 1, 0),
         discharge_info = 
           case_when(
             multiple_discharge==0 & expiration == 1 ~ "Expiration",
             multiple_discharge==0 & acute_rehabilitation == 1 ~ "Acute rehabilitation",
             multiple_discharge==0 & subacute_rehabilitation == 1 ~ "Subacute rehabilitation",
             multiple_discharge==0 & inpatient_rehabilitation == 1 ~ "Inpatient rehabilitation",
             multiple_discharge==0 & assited_living_facility == 1 ~ "Assited living facility",
             multiple_discharge==0 & home_health == 1 ~ "Home health",
             multiple_discharge==0 & home_with_no_needs == 1 ~ "Home with no needs",
             multiple_discharge==0 & hospice == 1 ~ "Hospice",
             multiple_discharge==0 & other_discharge_location == 1 ~ "Other discharge location",
             multiple_discharge==1 ~ "multiple discharge",
             TRUE ~ "other/unknown"),
         discharge_nihss=as.numeric(discharge_nihss))

table1(~ admission_nihss+discharge_nihss|discharge_info,overall = c(left = "Overall"),data=test)

table(stroke$expiration,useNA = "ifany")
```

Everyone (9 patients) with a sum of two for discharge outcomes have home_health and hospice. Only one patient (119) has acute_rehabilitation and home_health.

For the discharge, make a case_when to include all different levels. starts at discharge (fq) to other discharge location (ga). Expiration is if they died before discharge. Check if they have a yes and a discharge location (quality conflict)

```{r}
table(is.na(stroke$pt_am_pac_score))
table(is.na(stroke$admission_mrs_score))
table(is.na(stroke$admission_nihss))

stroke |> 
  filter(is.na(pt_am_pac_score),is.na(admission_mrs_score),is.na(admission_nihss)) |>
  select(id_number,date_of_stroke)
```


Everyone (9 patients) with a sum of two for discharge outcomes have home_health and hospice. Only one patient (119) has acute_rehabilitation and home_health.

For the discharge, make a case_when to include all different levels. starts at discharge (fq) to other discharge location (ga). Expiration is if they died before discharge. Check if they have a yes and a discharge location (quality conflict)

look at insulin as well, medication notes

last known well time (LKWT)

```{r Relationship between outcome variables}
stroke |> 
  ggplot(aes(x=admission_nihss,y=discharge_nihss))+
  geom_point()
```


```{r}
stroke
```


```{r inital model}
mdl1<-lm(admission_nihss~location_of_stroke+prior_stroke+covid_19+race+sex+age+
           bmi_kg_m_2+cholesterol_mg_d_l+dm+depression+length_of_admission_days,
         data=stroke)
summary(mdl1)
plot(mdl1)
```

```{r}
hist(stroke$admission_nihss)
hist(as.numeric(stroke$discharge_mrs_score))

table(stroke$discharge_mrs_score,useNA="ifany")

stroke$ot_am_pac_score
table(stroke$pt_am_pac_score,useNA="ifany")
table(stroke$ot_am_pac_score,useNA="ifany")
```

