---
title: "Final Regression Analysis"
author: "Lidio Jaimes Jr"
date: "2025-04-13"
output: pdf_document
---
```{r Importing Libraries}
library(naniar)
library(forestplot)
library(mice)
# library(visdat)
# library(VIM)
# library(corrplot)
# library(kableExtra)
```

# Multiple Imputation
This data set will exclude **statin at arrival**. If we want to include it, then look after second year of data entry.

## Importing the data
```{r importing dataset for this mdl1}
mdl1_data <- 
  stroke |> 
  mutate(
    caregiver = case_when(
      caregiver == "Present" ~ 1,
      caregiver == "Absent" ~ 0),
    prior_stroke = case_when(
      prior_stroke == "No" ~ 0,
      prior_stroke == "Yes" ~ 1),
    sex_male = ifelse(sex == "Male", 1, 0),
    dm = ifelse(dm == "No", 0, 1),
    id_number = as.numeric(id_number),
    location_of_stroke = 
      factor(location_of_stroke,
             levels = c("Cortical","Multiple Locations",
                        "Subcortical","Cerebellum",
                        "Brainstem")),
    side_of_stroke = factor(side_of_stroke, 
                            levels = c("Left","Right","Bilateral")),
    admission_unit = factor(admission_unit,
                            levels = c("Admission transfer from ED", "Observation Unit",
                                       "Stroke Unit", "ICU", "Other"))) |> 
  filter(length_of_admission_days!=0) |> #log gives us -Inf, also removes NA's
  select(#id_number, location_of_stroke,
         admission_nihss, admission_mrs_score, admission_unit,
         caregiver, prior_stroke, race, age, sex_male, 
         bmi_kg_m_2, cholesterol_mg_d_l, dm, antidep_arrival, anxio_arrival,
         htn, ischemic_heart_disease, cigarette_smoking, lesion_cortical, 
         length_of_admission_days, side_of_stroke)

mdl1_data$caregiver <- factor(mdl1_data$caregiver)
mdl1_data$prior_stroke <- factor(mdl1_data$prior_stroke)
mdl1_data$antidep_arrival <- factor(mdl1_data$antidep_arrival)
mdl1_data$anxio_arrival <- factor(mdl1_data$anxio_arrival)
mdl1_data$lesion_cortical <- factor(mdl1_data$lesion_cortical)
mdl1_data$sex_male <- factor(mdl1_data$sex_male)
mdl1_data$dm <- factor(mdl1_data$dm)
mdl1_data$htn <- factor(mdl1_data$htn)
mdl1_data$ischemic_heart_disease <- factor(mdl1_data$ischemic_heart_disease)

mdl1_data_complete <-
  mdl1_data[complete.cases(mdl1_data),]
```

## Missingness
```{r Visualizing missing data}
visdat::vis_miss(select_if(mdl1_data,~ any(is.na(.)))) +
  labs(title="Missingness in stroke data")

VIM::aggr(mdl1_data,
          col = c("blue","red"),
          numbers = TRUE,
          labels = names(mdl1_data),
          cex.axis=.4, gap=3,
          ylab = c("Histogram of missing data", "Pattern"))
```

## Checking conditions for MICE

```{r MCAR vs MAR}
mcar_test(mdl1_data[mdl1_data$length_of_admission_days!=0,]) #NOT MCAR
miss_var_summary(mdl1_data)
```

Assuming the data is MAR. **What drives the missingness for admission scores?**

This is just to check before runnning MICE
```{r Normality, eval=FALSE}
mdl1_data_numeric <- mdl1_data |> select(-1) |> select_if(is.numeric)
for (col in colnames(mdl1_data_numeric)) {
  hist(mdl1_data_numeric[[col]], main = paste(col))
}
```

Only age, bmi, cholesterol, and log length hospital stay can be argued for being normally distributed

```{r Multicollinearity}
mdl1_data |> 
  select_if(is.numeric) |> 
  na.omit() |> # 271 patients
  cor() |> 
  corrplot::corrplot.mixed()
```

No alarming multicollinearity

# Initial Model
$$\log[\text{Length of admission days}_i] = \beta_0 + \beta_1\text{AdmissionNIHSS}_i + \beta_2\text{AdmissionMRS}_i + \beta_3\text{Side of Stroke-Right}_i + \beta_4\text{Side of Stroke-Bilateral}_i +  \\ \beta_5\text{Cortical Lesion}_i + \beta_6\text{AdmissionUnit-Observation}_i + \beta_7\text{AdmissionUnit-Stroke}_i +\beta_8\text{AdmissionUnit-ICU}_i + \\ \beta_9\text{AdmissionUnit-Other}_i +  \beta_{10}\text{Prior Stroke}_i + \beta_{11}\text{Diabetes}_i + \beta_{12}\text{Hypertension}_i + \beta_{13}\text{IschemicHeartDisease}_i + \\ \beta_{14}\text{Antidepressants}_i + \beta_{15}\text{Anxiolytics}_i + \beta_{16}\text{SmokingHistory-Former}_i + \beta_{17}\text{SmokingHistory-Current}_i + \\ \beta_{18}\text{CaregiverPresence}_i +  \beta_{19}\text{Race-Black}_i + \beta_{20}\text{Race-Other}_i + \beta_{21}\text{Age}_i + \beta_{22}\text{Sex-Male}_i + \beta_{23}\text{BMI}_i + \beta_{24}\text{Cholesterol}_i + \epsilon_i$$

## Running MICE and Regression
```{r MICE}
mdl1_data_imp <- mdl1_data 
imp <- mice(mdl1_data_imp, maxit=0)
predM <- imp$predictorMatrix
meth <- imp$method

predM[, c("length_of_admission_days")] <- 0

set.seed(5010) 
imp2 <- mice(mdl1_data_imp, m = 100, 
             predictorMatrix = predM, 
             method = meth, print =  FALSE)

imp_long <- mice::complete(imp2, action="long", include = TRUE)
imp_long_mids<-as.mids(imp_long)
fitimp <- with(imp_long_mids,
               lm(log(length_of_admission_days) ~ admission_nihss + admission_mrs_score + 
                    side_of_stroke + lesion_cortical + admission_unit + prior_stroke + 
                    dm + htn + ischemic_heart_disease + antidep_arrival + anxio_arrival +
                    cigarette_smoking + caregiver + race + age + sex_male + bmi_kg_m_2 +
                    cholesterol_mg_d_l))
```

## MICE output
```{r MICE output}
res <- pool(fitimp)
summary(res)

regression_table <- as.data.frame(summary(res))
regression_table[,2:6] <- round(regression_table[,2:6],5)
regression_table$estimate <- exp(regression_table$estimate)
regression_table$term <- c("Intercept", "Admission NIHSS", "Admission MRS", "Side of Stroke - Right", 
                           "Side of Stroke - Bilateral", "Cortical Lesion", "Admission Unit - Observation", 
                           "Admission Unit - Stroke", "Admission Unit - ICU", "Admission Unit - Other", 
                           "Prior Stroke", "Diabetes", "Hypertension", "Ischemic Heart Disease", 
                           "Antidepressants at Arrival", "Anxiolytics at Arrival", "SmokingHistory - Former",
                           "SmokingHistory - Current", "Caregiver Presence", "Race - Black", "Race - Other", 
                           "Age", "Sex - Male", "BMI", "Cholesterol")
colnames(regression_table) <- c("Variable", "Estimate", "Standard Error", "Statistic", "Degrees of Freedom", "p-value")
regression_table$`p-value` <- ifelse(regression_table$`p-value` < 0.001,
                                     "< 0.001", 
                                     round(regression_table$`p-value`, 3))

kableExtra::kable(regression_table)
glance(res)
```

# Stepwise
```{r MICE Stepwise}
set.seed(5010)
selected_vars <- list()
for (i in 1:100) {
  data_i <- complete(imp2, i)
  full_model <- lm(log(length_of_admission_days) ~ admission_nihss + admission_mrs_score + 
                    side_of_stroke + lesion_cortical + admission_unit + prior_stroke + 
                    dm + htn + ischemic_heart_disease + antidep_arrival + anxio_arrival +
                    cigarette_smoking + caregiver + race + age + sex_male + bmi_kg_m_2 +
                    cholesterol_mg_d_l,
                   data = data_i)
  step_model <- step(full_model, direction = "both", trace = FALSE) #stepwise on each mdl
  selected_vars[[i]] <- names(coef(step_model))
}

sort(table(unlist(selected_vars)), decreasing = TRUE) #frequency of selected variables, want 70%

fit_final <- with(imp2, #run final stepwise mdl to all 100 datasets
                  lm(log(length_of_admission_days) ~ 
                       admission_nihss + admission_mrs_score + side_of_stroke + 
                       admission_unit + htn + age + bmi_kg_m_2))
```

$$\log[\text{Length of admission days}_i] = \beta_0 + \beta_1\text{AdmissionNIHSS}_i + \beta_2\text{AdmissionMRS}_i + \beta_3\text{Side of Stroke-Right}_i + \beta_4\text{Side of Stroke-Bilateral}_i +  \\ \beta_5\text{AdmissionUnit-Observation}_i + \beta_6\text{AdmissionUnit-Stroke}_i +\beta_7\text{AdmissionUnit-ICU}_i +  \beta_8\text{AdmissionUnit-Other}_i + \\\beta_{9}\text{Hypertension}_i + \beta_{10}\text{Age}_i + \beta_{11}\text{BMI}_i + \epsilon_i$$

```{r Stepwise Results}
res_final <- pool(fit_final)
summary(res_final)

regression_table <- as.data.frame(summary(res_final))
regression_table[,2:6] <- round(regression_table[,2:6],5)
regression_table$estimate <- exp(regression_table$estimate)
regression_table$term <- c("Intercept", "Admission NIHSS", "Admission MRS", "Side of Stroke - Right", 
                           "Side of Stroke - Bilateral", "Admission Unit - Observation", "Admission Unit - Stroke", 
                           "Admission Unit - ICU", "Admission Unit - Other", "Hypertension", "Age", "BMI")
colnames(regression_table) <- c("Variable", "Estimate", "Standard Error", "Statistic", "Degrees of Freedom", "p-value")
regression_table$`p-value` <- ifelse(regression_table$`p-value` < 0.001,
                                     "< 0.001", 
                                     round(regression_table$`p-value`, 3))

kableExtra::kable(regression_table)
glance(res_final)
```

## Forest plot for stepwise model
```{r Making Forest Plot Dataset}
fp_data2 <- regression_table[-1,]
fp_data2$labeltext <- c("Admission NIHSS", "Admission MRS", "Side of Stroke - Right", 
                       "Side of Stroke - Bilateral", "Admission Unit - Observation", 
                       "Admission Unit - Stroke", "Admission Unit - ICU", "Admission Unit - Other", 
                       "Hypertension", "Age", "BMI")
fp_data2$display_est <- round(fp_data2$Estimate,3)
fp_data2$lower<- round(fp_data2$Estimate - 1.96 * fp_data2$`Standard Error`,3)
fp_data2$upper<- round(fp_data2$Estimate + 1.96 * fp_data2$`Standard Error`,3)
fp_data2$ci <- paste0("(", fp_data2$lower, ", ", fp_data2$upper, ")")
fp_data2$pval <- fp_data2$`p-value`

table_text <- cbind(
  c(fp_data2$labeltext),
  c(as.character(fp_data2$display_est)),
  c(fp_data2$ci),
  c(as.character(fp_data2$pval)))
```

```{r Forest Plot}
forestplot(labeltext = table_text,
           mean = c(fp_data2$display_est),
           lower = c(fp_data2$lower),
           upper = c(fp_data2$upper),
           zero = 1,
           boxsize = 0.3,
           lineheight = unit(1, "cm"),
           col = forestplot::fpColors(box="#BA0C2F", line="black", summary="royalblue"),
           xlab = "",
           title = "Forest Plot of Regression Coefficients of Stepwise Model",
           txt_gp = fpTxtGp(
             label = gpar(fontfamily = "Times", col = "black"),
             ticks = gpar(fontfamily = "Times", col = "black")))|> 
  fp_add_header("Covariates" |> fp_align_center(),
                    "Est." |> fp_align_center(),
                    "95% C.I." |> fp_align_center(),
                    "P-Val" |> fp_align_center()) |> 
  fp_set_zebra_style("#EFEFEF")
```

# interest in interaction terms 
running stepwise on 100 imputed datasets with all interaction terms
```{r}
set.seed(5010)
selected_vars <- list()
for (i in 1:100) {
  data_i <- complete(imp2, i)
  
  full_model <- lm(log(length_of_admission_days) ~ .^2,
                   data = data_i)
  step_model <- step(full_model, direction = "both", trace = FALSE)
  selected_vars[[i]] <- names(coef(step_model))
}

var_freq <- table(unlist(selected_vars))
var_freq <- sort(var_freq, decreasing = TRUE)
print(var_freq)

final_vars <- names(var_freq[var_freq >= 70])
final_vars <- final_vars[final_vars != "(Intercept)"]
print(final_vars)
```

