---
title: "Methodology"
author: "Lidio Jaimes Jr"
date: "2025-02-21"
output: pdf_document
---
```{r,include=FALSE}
library(tidyverse)
library(forcats)
library(table1)
```

# Dan and Christina Functions 
```{r,include=FALSE}
##
#####	Collection of functions written by D. Gillen that are used
#####	throughout Stat 211
#####
#####	Author: D. Gillen
#####	Course: Stat 211, Winter 2022
##
##
#####	Helper function to return vector for binary test
##
ifelse1 <- function (test, yes, no){
    if (test) yes
    else no
}

##
#####  Function to produce CIs for LM parameters (or exponentiated parameters)
##
lmCI <- function( model, expcoef=FALSE, robust=FALSE ){
	coef <- summary( model )$coef[,1]
	se <- ifelse1( robust, robust.se.lm(model)[,2], summary( model )$coef[,2] )
	tvalue <- coef / se
	pvalue <- 2*(1-pt(abs(tvalue), model$df.residual))
	if( expcoef ){
		ci95.lo <- exp( coef - qt(.975, model$df.residual) * se )
		ci95.hi <- exp( coef + qt(.975, model$df.residual) * se )
		est <- exp( coef )
	}
	else{
		ci95.lo <- coef - qt(.975, model$df.residual) * se
		ci95.hi <- coef + qt(.975, model$df.residual) * se
		est <- coef
	}
	rslt <- round( cbind( est, ci95.lo, ci95.hi, tvalue, pvalue ), 4 )
	colnames( rslt ) <- ifelse1( 	robust, 	
					c("Est", "robust ci95.lo", "robust ci95.hi", "robust t value", "robust Pr(>|t|)"),
					c("Est", "ci95.lo", "ci95.hi", "t value", "Pr(>|t|)") )			
	colnames( rslt )[1] <- ifelse( expcoef, "exp( Est )", "Est" )
	rslt
	}


##
#####	Function to plot dfBetas resulting from a lm() object
##
plot.dfbeta <- function( dfbeta.fit, labels ){
	oldmar <- par()$mar
	par( mar=c(5, 4, 3+.75*dim(dfbeta.fit)[2], 2) + 0.1 )
	plot( c(1,dim(dfbeta.fit)[1]*1.1), range(dfbeta.fit)*1.1, xlab="Observation", ylab="dfBeta", type="n" )
	for( i in 2:dim(dfbeta.fit)[2] ){
			points( 1:dim(dfbeta.fit)[1], dfbeta.fit[,i], col=i )
			text( 1:dim(dfbeta.fit)[1]+1, dfbeta.fit[,i]+.1, labels=labels, col=i )
			mtext( colnames( dfbeta.fit )[i], col=i, line=-1+i )
	}  
	abline( h=c(-1,1)*(2/sqrt(dim(dfbeta.fit)[1])), col="red", lwd=2 )
	par( mar=oldmar )
	}
	

##
##### 	Compute robust (sandwich) variance-covariance estimate for a LM
##
robust.vcov.lm <- function( lm.obj ){
  X <- model.matrix( lm.obj )
  eps <- lm.obj$residuals
  robust.cov <- solve( t(X)%*%X ) %*%( t(X) %*% diag(eps^2) %*% X ) %*% solve( t(X)%*%X )
  dimnames( robust.cov ) <- dimnames( vcov(lm.obj) )
  return( robust.cov )
}

##
#####
#####	robust.se.lm() is a function to compute the Huber-White sandwich variance estimator
#####	for the linear regression model
#####	
##
robust.se.lm <- function( model) { 
  s <- summary( model) 
  X <- model.matrix( model )
  sandwich.cov <- robust.vcov.lm( model )
  sand.se <- sqrt( diag( sandwich.cov )) 
  t <- model$coefficients/sand.se
  p <- 2*pt( -abs( t ), dim(X)[1]-dim(X)[2] ) 
  ci95.lo <- model$coefficients - qt( .975, dim(X)[1]-dim(X)[2] ) * sand.se
  ci95.hi <- model$coefficients + qt( .975, dim(X)[1]-dim(X)[2] ) * sand.se
  rslt <- cbind( model$coefficients, sand.se, ci95.lo, ci95.hi, t, p ) 
  dimnames(rslt)[[2]] <- c( dimnames( s$coefficients )[[2]][1], "Robust SE", "ci95.lo", "ci95.hi", dimnames( s$coefficients )[[2]][3:4] ) 
  rslt 
} 


##
#####  Function to exponentiate coefficients and produces CIs for GLMs
##
glmCI <- function( model, transform=TRUE, robust=FALSE ){
	link <- model$family$link
	coef <- summary( model )$coef[,1]
	se <- ifelse1( robust, robust.se.glm(model)[,2], summary( model )$coef[,2] )
	zvalue <- coef / se
	pvalue <- 2*(1-pnorm(abs(zvalue)))

	if( transform & is.element(link, c("logit","log")) ){
		ci95.lo <- exp( coef - qnorm(.975) * se )
		ci95.hi <- exp( coef + qnorm(.975) * se )
		est <- exp( coef )
	}
	else{
		ci95.lo <- coef - qnorm(.975) * se
		ci95.hi <- coef + qnorm(.975) * se
		est <- coef
	}
	rslt <- round( cbind( est, ci95.lo, ci95.hi, zvalue, pvalue ), 4 )
	colnames( rslt ) <- ifelse1( 	robust, 	
					c("Est", "robust ci95.lo", "robust ci95.hi", "robust z value", "robust Pr(>|z|)"),
					c("Est", "ci95.lo", "ci95.hi", "z value", "Pr(>|z|)") )			
	colnames( rslt )[1] <- ifelse( transform & is.element(link, c("logit","log")), "exp( Est )", "Est" )
	rslt
	}


##
#####  Function to collapse grouped binary data to binomial
##
collapse <- function( data, outcome ){
	index <- (1:length(names(data)))[ names(data)==outcome ]
	y <- data[,index]
	covnames <- names( data )[-index]
	data <- data[,-index]
	if( is.null( dim( data ) ) ){
		rslt <- aggregate( y, list(data), FUN=length)
		rslt <- as.data.frame( cbind( rslt, aggregate( y, list(data), FUN=sum)[dim(rslt)[2]] ) )
	}
	else{
		rslt <- aggregate( y, data, FUN=length)
		rslt <- as.data.frame( cbind( rslt, aggregate( y, data, FUN=sum)[dim(rslt)[2]] ) )
	}
	names( rslt ) <- c( covnames, "n", paste("n.", outcome, sep="") )
	rslt
}


##
#####	Function to compute deviance (LR) test p-Value
##
lrtest <- function( fit1, fit2 ){
	cat( "\nAssumption: Model 1 nested within Model 2\n\n" )
	rslt <- anova( fit1, fit2 )
	rslt <- cbind( rslt, c("", round( pchisq( rslt[2,4], rslt[2,3], lower.tail=FALSE ), 4 ) ) )
	rslt[,2] <- round( rslt[,2], 3 )
	rslt[,4] <- round( rslt[,4], 3 )
	rslt[1,3:4] <- c( "", "" )
	names( rslt )[5] <- "pValue"
	rslt
}


##
#####	H-L goodness of fit test
##
binary.gof <- function( fit, ngrp=10, print.table=TRUE ){
	y <- fit$y
	phat <- fitted( fit )
	fittedgrps <- cut( phat, quantile( phat, seq(0,1,by=1/ngrp) ), include.lowest=TRUE )
	n <- aggregate( y, list( fittedgrps ), FUN=length )[,2]
	Obs <- aggregate( y, list( fittedgrps ), FUN=sum )[,2]
	Exp <- aggregate( phat, list( fittedgrps ), FUN=sum )[,2]
	if( print.table==TRUE ){
		cat( "\nFitted Probability Table:\n\n" )
		rslt <- as.data.frame( cbind( 1:ngrp, n, Obs, Exp ) )
		names( rslt )[1] <- "group"
		print( rslt )
	}
	chisqstat <- sum( (Obs - Exp)^2 / ( Exp*(1-Exp/n) ) )
	df <- ngrp - 2
	pVal <- pchisq( chisqstat, df, lower.tail=FALSE )
	cat( "\n Hosmer-Lemeshow GOF Test:\n\n" )
	cbind( chisqstat, df, pVal )
}

##
##### 	Compute robust (sandwich) variance-covariance estimate for a GLM
##
robust.vcov.glm <- function(glm.obj){
	if (is.matrix(glm.obj$x)) 
		xmat<-glm.obj$x
	else {
		mf<-model.frame(glm.obj)
		xmat<-model.matrix(terms(glm.obj),mf)		
	}
	umat <- residuals(glm.obj,"working")*glm.obj$weights*xmat
	modelv<-summary(glm.obj)$cov.unscaled
	robust.cov <- modelv%*%(t(umat)%*%umat)%*%modelv
	dimnames( robust.cov ) <- dimnames( vcov(glm.obj) )
	return( robust.cov )
}
	
	
##
#####	Function to compute robust se for glms
##
robust.se.glm <- function(glm.obj){
	## 	Compute robust (sandwich) variance estimate
	robust.cov <- robust.vcov.glm(glm.obj)
	
	##	Format the model output with p-values and CIs
	s <- summary( glm.obj) 
	robust.se <- sqrt( diag( robust.cov )) 
	z <- glm.obj$coefficients/robust.se
	p <- 2*pnorm( -abs( z ) ) 
	ci95.lo <- glm.obj$coefficients - qnorm( .975 ) * robust.se
	ci95.hi <- glm.obj$coefficients + qnorm( .975 ) * robust.se
	rslt <- cbind( glm.obj$coefficients, robust.se, ci95.lo, ci95.hi, z, p ) 
	dimnames(rslt)[[2]] <- c( dimnames( s$coefficients )[[2]][1], "Robust SE", "ci95.lo", "ci95.hi", "z value", "Pr(>|z|)" ) 
	rslt 
	}

##
#####	Function to summarize the multinomial fit
##
summ.mfit <- function( model ){
	s <- summary( model )
	for( i in 1:length(model$coef) ){
		cat( "\nLevel ", model$lev[i+1],  "vs. Level ", model$lev[1], "\n" )
		coef <- s$coefficients[i,]
		rrr <- exp( coef )
		se <- s$standard.errors[i,]
		zStat <- coef / se
		pVal <- 2*pnorm( abs(zStat), lower.tail=FALSE )
		ci95.lo <- exp( coef - qnorm(.975)*se )
		ci95.hi <- exp( coef + qnorm(.975)*se )
		rslt <- cbind( rrr, se, zStat, pVal, ci95.lo, ci95.hi )
		print( round( rslt, 3 ) )
	}
}

#
##
#####	Function to summarize a proportional odds fit
##
olrCI <- function( model, transform=TRUE ){
		coef <- summary( model )$coef[ 1:length(model$coef), ]
		if( is.null( dim( coef ) ) ) coef <- t( as.matrix(coef) )
		zStat <- as.matrix(coef[,3])
		pVal <- 2*pnorm( abs(zStat),lower.tail=FALSE )
		if( transform ){
			ci95.lo <- exp( coef[,1] - qnorm(.975) * coef[,2] )
			ci95.hi <- exp( coef[,1] + qnorm(.975) * coef[,2] )
			est <- exp( coef[,1] )
		}
		else{
			ci95.lo <- coef[,1] - qnorm(.975) * coef[,2]
			ci95.hi <- coef[,1] + qnorm(.975) * coef[,2]
			est <- coef[,1]
		}
		rslt <- round( cbind( est, ci95.lo, ci95.hi, zStat, pVal), 4 )
		if( transform ) colnames( rslt ) <- c( "exp( Est )", "ci95.lo", "ci95.hi", "z value", "Pr(>|z|)" )
			else colnames( rslt ) <- c( "est", "ci95.lo", "ci95.hi", "z value", "Pr(>|z|)" )
		rslt
}


mvWald <- function(model, cov.names, robust=FALSE) {
  p <- length(model$coef)
  r <- length(cov.names)
  names.model <- names(model$coef)
  new.index <- c((1:p)[names.model %in% cov.names], (1:p)[!names.model %in%
                                                            cov.names])
  beta <- model$coef[new.index]
  if (robust)
    cov.beta <- robust.vcov.glm(model) else cov.beta <- vcov(model)
  Info <- solve(cov.beta[new.index, new.index])
  Info.11 <- Info[1:r, 1:r]
  Info.12 <- Info[1:r, (r + 1):p]
  Info.22 <- Info[(r + 1):p, (r + 1):p]
  Info.11.2 <- Info.11 - Info.12 %*% solve(Info.22) %*% t(Info.12)
  chiStat <- t(beta[1:r]) %*% Info.11.2 %*% beta[1:r]
  pVal <- 1 - pchisq(chiStat, df = r)
  rslt <- as.data.frame(cbind(round(chiStat, 2), round(r), round(pVal,3)))
  dimnames(rslt)[[2]] <- c("ChiStat", "DF", "pValue")
  cat("\nTest of H_0: ")
  for (i in 1:r) {
    if (i < r)
      cat(cov.names[i], " = ", sep = "") else cat(cov.names[i], " = 0 :\n\n", sep = "")
  }
  rslt
}

##  Example use
#mvWald(fit.1, cov.names = c("factor(k5)1", "factor(k5)2"), robust=TRUE )




linContr.glm <- function(contr.names, contr.coef = rep(1, length(contr.names)),
                         model, transform = TRUE, robust = FALSE) {
  beta.hat <- model$coef
  if (robust)
    cov.beta <- robust.vcov.glm(model) else cov.beta <- vcov(model)
  contr.index <- match(contr.names, dimnames(cov.beta)[[1]])
  beta.hat <- beta.hat[contr.index]
  cov.beta <- cov.beta[contr.index, contr.index]
  est <- contr.coef %*% beta.hat
  se.est <- sqrt(contr.coef %*% cov.beta %*% contr.coef)
  zStat <- est/se.est
  pVal <- 2 * pnorm(abs(zStat), lower.tail = FALSE)
  ci95.lo <- est - qnorm(0.975) * se.est
  ci95.hi <- est + qnorm(0.975) * se.est
  link <- model$family$link
  if (transform & is.element(link, c("logit", "log"))) {
    ci95.lo <- exp(ci95.lo)
    ci95.hi <- exp(ci95.hi)
    est <- exp(est)
    cat("\nTest of H_0: exp( ")
    for (i in 1:(length(contr.names) - 1)) {
      cat(contr.coef[i], "*", contr.names[i], " + ", sep = "")
    }
    cat(contr.coef[i + 1], "*", contr.names[i + 1], " ) = 1 :\n\n",
        sep = "")
  } else {
    cat("\nTest of H_0: ")
    for (i in 1:(length(contr.names) - 1)) {
      cat(contr.coef[i], "*", contr.names[i], " + ", sep = "")
    }
    cat(contr.coef[i + 1], "*", contr.names[i + 1], " = 0 :\n\n", sep = "")
  }
  rslt <- data.frame(est, se.est, zStat, pVal, ci95.lo, ci95.hi)
  colnames(rslt)[1] <- ifelse(transform && is.element(link, c("logit",
                                                              "log")), "exp( Est )", "Est")
  round(rslt, 3)
}
```

```{r regression tables,include=FALSE}
format_lm_output <- function(data, sigfigs = 2, robust = TRUE){
  ########## round significant figures 
  ########## use the format function with nsmall and digits = 2
  ########## exp( Est ) ci95.lo ci95.hi  z value Pr(>|z|)
  #keep <- c("exp( Est )", "ci95.lo", "ci95.hi", "Pr(>|z|)")
  # Est ci95.lo ci95.hi t value Pr(>|t|)
  
  if(robust){
    ### "Est", "robust ci95.lo", "robust ci95.hi", "robust z value", "robust Pr(>|z|)"
    keep <- c("Est", "robust ci95.lo", "robust ci95.hi", "robust Pr(>|t|)")
  } else {
    keep <- c("Est", "ci95.lo", "ci95.hi", "Pr(>|t|)")
  }
  ### if you are using a robust variance your arguments would be 
  
  output <- data[ , colnames(data) %in% keep]
  colnames(output) <- c("Est", "ci95.lo", "ci95.hi", "pval")
  output <- as.data.frame(output)
  
  output <- cbind(format(output$Est, nsmall = sigfigs , digits = sigfigs, scientific = FALSE )
                  , paste("(", format(output$ci95.lo, nsmall = sigfigs , digits = sigfigs, scientific = FALSE ), ",", format(output$ci95.hi, nsmall = sigfigs , digits = sigfigs, scientific = FALSE ),")")
                  , format(output$pval, nsmall = sigfigs, digits = sigfigs, scientific = FALSE)
  )
  
  colnames(output) <- c("Est", "95 C.I.", "P-Val")
  rownames(output) <- rownames(data)
  
  return(output)
  
}


table1_descriptive <- function(data, vars, group, categorical){
  # , sigdigits=2 may want to add this argument in later 
  
  n.var <- length(vars)
  
  groupcounts <- table(data[ , colnames(data) %in% group])
  
  groupNum <- paste("N = ", groupcounts)
  table_1 <- rbind(rownames(groupcounts), groupNum)
  
  for (i in 1:n.var) {
    
    if(vars[i] %in% categorical){
      
      counts <- table(data[ , colnames(data) %in% vars[i] ], data[, colnames(data) %in% group], useNA = "ifany")
      
      freq <- counts / matrix(rep(groupcounts, each = dim(counts)[1]), nrow = dim(counts)[1])
      freq <- format(freq, nsmall = 1, digits = 1)
      counts <- format(counts, nsmall = 2, digits = 2)
      
      temp <- matrix(paste(counts, "(", freq, ")"), nrow = dim(counts)[1]) # will have to format significant figures 
      rownames(counts) <- ifelse(is.na(rownames(counts)), "Missing", rownames(counts))
      rownames(temp) <- paste("    ", rownames(counts)) #for excel table 
      #rownames(temp) <- ifelse(is.na(rownames(temp)), "Missing", rownames(temp))
      #rownames(temp) <- paste("\\indent ", rownames(counts))
      #rownames(temp) <- paste("\\hspace{3mm} ", rownames(counts))
      #colnames(temp) <- colnames(counts)
      #colnames(temp) <- ifelse(is.na(colnames(temp)), "Missing", colnames(temp))
      
      filler <- matrix(rep("", length(groupcounts)), nrow = 1)
      rownames(filler) <- vars[i]
      colnames(filler) <- colnames(counts)
      
      table_1 <- rbind(table_1, filler, temp) 
      # appending to table 1
      
      
    } else{
      
      ########## mean 
      means_group <- tapply(data[ , colnames(data) %in% vars[i] ], data[, colnames(data) %in% group], mean, na.rm=TRUE)
      
      ########### formating sig figs
      means_group <- format(means_group, nsmall = 2, digits = 2)
      
      
      ############ sd 
      sds_group <- tapply(data[ , colnames(data) %in% vars[i] ], data[, colnames(data) %in% group], sd, na.rm=TRUE)
      
      
      ########### formating sig figs
      sds_group  <- format(sds_group, nsmall = 2, digits = 2)
      
      
      
      temp <- matrix(paste(means_group, "(", sds_group, ")"), nrow = 1) 
      
      colnames(temp) <- rownames(means_group)
      rownames(temp) <- vars[i]
      
      table_1 <- rbind(table_1, temp) 
      
      
    }
    
    
  }
  
  return(table_1)
}

dfbeta_lm_plots <- function(fit_model){

  deltabetas_model <- dfbeta(fit_model)
  
  for (i in 1:dim(deltabetas_model)[2]) {
    plot(seq(1:dim(deltabetas_model)[1]), deltabetas_model[, i], main = colnames(deltabetas_model)[i])
    
    # Add observation numbers as labels
    text(
      x = rownames(deltabetas_model), 
      y = deltabetas_model[, i], 
      labels = rownames(deltabetas_model), 
      pos = 4,  # Position labels to the right of the points
      cex = 0.7, # Size of text
      col = "red"
    )
  }
}

format_geeConstrast <- function(data_t){
  
  
  output <- cbind(
    data_t$Est
    , paste("(", data_t$ci95.lo, ",", data_t$ci95.hi, ")")
    , data_t$pVal)
  return(output)
  
}

dfbeta_geeglm_plots <- function(fit_model, data, family = gaussian, corstr = "ar1"){
  ######### library(geepack) using this function 
  #fit_model <- modelBaseline
  # data = ds_long_base
  
  gee_formula <- formula(fit_model)
  
  x_design <- as.data.frame(model.matrix(fit_model))
  subject_id <- data$subject_id[rownames(data) %in% as.numeric(rownames(x_design))]
  #y_vector <- model.response(model.frame(fit_model))
  #data_model <- as.data.frame(cbind(y_vector, x_design, subject_id))
  
  sub_unique <- unique(subject_id)
  n_unique <- length(sub_unique)
  
  beta0 <- coefficients(fit_model)
  
  
  delta_betas <- matrix(NA, ncol = dim(x_design)[2], nrow = n_unique)
  colnames(delta_betas) <- colnames(x_design)
  rownames(delta_betas) <- sub_unique 
  
  for (i in 1:n_unique) {
    delta_betas[i, ] <- coefficients(geeglm( gee_formula
                                             ,id = subject_id,
                                             data = data[data$subject_id != sub_unique[i],] ,
                                             ,family = family
                                             ,na.action = na.omit
                                             ,corstr = corstr)) - beta0
  }
  
  
  for (i in 1:dim(delta_betas)[2]) {
    plot(seq(1:dim(delta_betas)[1]), delta_betas[, i], main = colnames(delta_betas)[i])
    
    # Add observation numbers as labels
    text(
      x = seq(1:dim(delta_betas)[1]), 
      y = delta_betas[, i], 
      labels = unique(subject_id), 
      pos = 4,  # Position labels to the right of the points
      cex = 0.7, # Size of text
      col = "red"
    )
  }
  
  return(list(deltabetas = delta_betas))
}
```

# Reading in data for knit
```{r Reading in Data,include=FALSE}
stroke_raw <- 
  readxl::read_xlsx("Capstone Combined Data Collection Sheets - Summer 2024.xlsx", 
                    skip = 1,
                    na=c("", " ", "UNK", "Unk", "NA", "na", "Na", 9999, "????")) |> 
  janitor::clean_names()
stroke <- stroke_raw[-c(1,376), ] #get rid of dictionary and black row

# E336 (patient 333) and E340 (patient 337) have full date instead of just year
stroke[stroke$id_number == 333, "year_of_birth"] = 1905 #set as 1981 instead of 1905
stroke[stroke$id_number == 337, "year_of_birth"] = 1942
stroke$date_identified = openxlsx::convertToDate(stroke$date_identified)
stroke$date_of_stroke = openxlsx::convertToDate(stroke$date_of_stroke)
stroke$expiration_date_in_hospital = openxlsx::convertToDate(stroke$expiration_date_in_hospital)
stroke$expiration_date = openxlsx::convertToDate(stroke$expiration_date)
stroke$length_of_admission_days = round(as.numeric(stroke$length_of_admission_days),0)

stroke <- 
  stroke |> 
  filter(!id_number%in%c(211, 391, 528, 529, 536, 537, 555, 558, 576, 582),
         !id_number%in%c(316,322),
         ischemic == 1) 
```

```{r Data Wrangling,include=FALSE}
stroke$brainstem_stroke_lcation <- as.numeric(stroke$brainstem_stroke_lcation)

stroke <-
  stroke |>
  mutate(multiple_locations = ifelse(rowSums(stroke[,106:110], na.rm = TRUE) > 1, 1, 0),
         location_of_stroke = 
           case_when(
             multiple_locations==0 & brainstem_stroke_lcation == 1 ~ "brainstem",
             multiple_locations==0 & cerebellum_stroke_location == 1 ~ "cerebellum",
             multiple_locations==0 & cortical_stroke_location == 1 ~ "cortical",
             multiple_locations==0 & subcortical_stroke_location == 1 ~ "subcortical",
             multiple_locations==1 ~ "multiple locations"),
         location_of_stroke = factor(location_of_stroke,
                                     levels = c("brainstem","cerebellum",
                                                "cortical","subcortical",
                                                "multiple locations"),
                                     labels = c("Brainstem","Cerebellum",
                                                "Cortical","Subcortical",
                                                "Multiple Locations")),
         race = ifelse(race=="Other","Other","White"),
         race = case_when(
           race == "White" ~ "White",
           str_detect(patient_identification_notes, "African") ~ "Black",
           TRUE ~ "Other"),
         race = factor(race, levels = c("White", "Black", "Other")),
         age = as.numeric(age),
         admission_nihss = as.numeric(admission_nihss),
         cholesterol_mg_d_l = as.numeric(cholesterol_mg_d_l),
         prior_stroke = case_when(
           prior_stroke == 0 ~ "No",
           prior_stroke == 1 ~ "Yes"),
         prior_tia = case_when(
           prior_tia == 0 ~ "No",
           prior_tia == 1 ~ "Yes"),
         cholesterol_group = case_when(
           cholesterol_mg_d_l < 200 ~ "Normal",
           cholesterol_mg_d_l >= 200 & cholesterol_mg_d_l <=239 ~ "Borderline high",
           cholesterol_mg_d_l >= 240 ~ "High"),
         cholesterol_group = factor(cholesterol_group,
                                    levels = c("Normal","Borderline high","High")),
         cigarette_smoking = case_when(
           cigarette_smoking == "Former" | cigarette_smoking == "former" ~ "Former",
           cigarette_smoking == "Never" | cigarette_smoking == "NEver" ~ "Never",
           cigarette_smoking == "Current" | cigarette_smoking == "Current\r\n\rCurrent" ~ "Current"),
         cigarette_smoking = factor(cigarette_smoking,
                                    levels = c("Never","Former","Current")),
         #ischemic_heart_disease = ifelse(ischemic_heart_disease == 0, "No", "Yes"),
         ischemic_heart_disease = as.numeric(ischemic_heart_disease),
         htn = as.numeric(htn),
         antidepressant_use = as.numeric(antidepressant_use),
         anxiolytic_use = as.numeric(anxiolytic_use),
         date_cov19 = ifelse(date_of_stroke>="2020-03-15","after covid","before covid"),
         covid_19 = ifelse(date_cov19=="before covid","No",covid_19),
         covid_19 = case_when(
           covid_19 == "No" ~ "No",
           covid_19 == "Yes (Once)" ~ "Yes",
           covid_19 == "positive" ~ "Yes",
           covid_19 == "negative" ~ "No"),
         antidep_arrival = ifelse(antidepressant_use == 1, 1, 0),
         anxio_arrival = ifelse(anxiolytic_use == 1, 1, 0),
         statin_arrival = 
           ifelse(str_detect(medication_list_at_arrival, "Statin"), 1, 0),
         # anticoag_arrival =
         #   ifelse(str_detect(medication_list_at_arrival, "Anticoagulants"), 1, 0),
         dm = factor(dm,levels=c(0,1),labels=c("No","Yes")),
         depression = factor(depression,levels=c(0,1),labels=c("No","Yes")),
         length_of_admission_days = as.numeric(length_of_admission_days),
         caregiver = case_when(
           presence_of_caregiver == "0" ~ "Absent",
           presence_of_caregiver == "1" ~ "Present",
           TRUE ~ "Unknown"),
         caregiver = factor(caregiver),
         admission_mrs_score = as.numeric(admission_mrs_score),
         pt_am_pac_score = as.numeric(pt_am_pac_score),
         discharge_nihss = as.numeric(discharge_nihss),
         discharge_mrs_score = as.numeric(discharge_mrs_score),
         ot_am_pac_score = as.numeric(ot_am_pac_score))

stroke[stroke$id_number==532,"race"] = "Other"
stroke[stroke$id_number==223,"race"] = "Other"
```



# Research Question

A predictive model in acute stroke patient's care incorporating pre-existing conditions (diabetes, hypertension, CAD, ischemic heart disease), lifestyle factors (smoking, obesity, hypercholesterolemia), socioeconomic, demographic factors, presence of caregiver and medication use (anticoagulants, statins, antidepressants, anxiolytics) will estimate patients' length of hospital stay at initial assessment.

$$\text{Length of admission days} = \beta_0 + \beta_1\text{AdmissionNIHSS}_i + \beta_2\text{AdmissionMRS}_i + \beta_3\text{StrokeLocation-MultipleLocations}_i + \\ \beta_4\text{StrokeLocation-Subcortical}_i +  \beta_5\text{StrokeLocation-Cerebellum}_i + \beta_6\text{StrokeLocation-Brainstem}_i + \\ \beta_7\text{CaregiverPresence}_i + \beta_8\text{PriorStroke}_i + \beta_9\text{Race-Black}_i + \beta_{10}\text{Race-Other}_i + \beta_{11}\text{Age}_i + \\ \beta_{12}\text{Sex}_i + \beta_{13}\text{BMI}_i + \beta_{14}\text{Cholesterol}_i + \beta_{15}\text{Diabetes} + \beta_{16}\text{Statins}_i + \beta_{17}\text{Antidepressants}_i + \\ \beta_{18}\text{Anxiolytics}_i + \beta_{19}\text{Hypertension} + \beta_{20}\text{IschemicHeartDisease} + \beta_{21}\text{SmokingHistory} + \epsilon_i$$

Shorter version:
$$E[\text{length of admission days}] = \beta_0 + \beta_1\text{AdmissionNIHSS}_i + \beta_2\text{AdmissionMRS}_i + \overrightarrow{\boldsymbol{\gamma}}^T\overrightarrow{\boldsymbol{z}_i}$$

What is the variable for CAD?

# Model 1 after second year of data entry (want to see meds at arrival)

```{r Subsetting complete}
mdl1_data <- 
  stroke |> 
  mutate(
    caregiver = case_when(
      caregiver == "Present" ~ 1,
      caregiver == "Absent" ~ 0),
    prior_stroke = case_when(
      prior_stroke == "No" ~ 0,
      prior_stroke == "Yes" ~ 1),
    sex_male = ifelse(sex == "Male", 1, 0),
    dm = ifelse(dm == "No", 0, 1),
    id_number = as.numeric(id_number),
    location_of_stroke = 
      factor(location_of_stroke,
             levels = c("Cortical","Multiple Locations",
                        "Subcortical","Cerebellum",
                        "Brainstem"))) |>
  filter(id_number>=375) |> #, #want everyone after second year of entry
  select(id_number,length_of_admission_days, admission_nihss, admission_mrs_score, 
         location_of_stroke, caregiver, prior_stroke, race, age, sex_male, 
         bmi_kg_m_2, cholesterol_mg_d_l, dm, antidep_arrival, anxio_arrival,
         statin_arrival, htn, ischemic_heart_disease)
```

```{r Full model base}
mdl1 <-
  lm(length_of_admission_days ~ admission_nihss + admission_mrs_score + 
       location_of_stroke + caregiver + prior_stroke + race + sex_male + 
       bmi_kg_m_2 + cholesterol_mg_d_l + dm + antidep_arrival + anxio_arrival +
       I((age - mean(age))/10) + statin_arrival + htn + ischemic_heart_disease, 
     data = mdl1_data_complete)
summary(mdl1)
plot(mdl1)
```

```{r Robust Variance Estimator for Full model}
#Run Dan211Functions First
library(kableExtra)
full_model_table <- lmCI(mdl1, robust = TRUE)
rownames(full_model_table) <- 
  c("Intercept", "Admission NIHSS", "Admission MRS", 
    "Diff b/w Multiple locations & Cortical", 
    "Diff b/w Subcortical & Cortical",
    "Diff b/w Cerebellum & Cortical",
    "Diff b/w Brainstem & Cortical",
    "Caregiver", "Prior Stroke", "Diff b/w Black & White", 
    "Diff b/w Other & White", "Diff b/w Male & Female",
    "BMI", "Cholesterol", "Diabetes", 
    "Antidepressants at Arrival", "Anxiolytic at Arrival", 
    "Centered Age per 10 years", "Statin at Arrival", "Hypertension", 
    "Ischemic Heart Disease")
full_final_table <- kable(format_lm_output(full_model_table), file = "html",
      caption = "Stroke Model Using All Covariates and Robust Variance Estimator")
full_final_table
```

```{r Stepwise}
#step(mdl1)
mdl1_step <- 
  lm(length_of_admission_days ~ admission_nihss + admission_mrs_score + 
       race + cholesterol_mg_d_l + I((age - mean(age))/10), 
     data = mdl1_data_complete)
summary(mdl1_step)
plot(mdl1_step)
```

```{r Robust Variance Estimator for Stepwise}
#Run Dan211Functions First
library(kableExtra)
step_model_table <- lmCI(mdl1_step, robust = TRUE)
rownames(step_model_table) <- 
  c("Intercept", "Admission NIHSS", "Admission MRS", 
    "Diff b/w Black & White", "Diff b/w Other & White", 
    "Cholesterol", "Centered Age per 10 years")
step_final_table <- kable(format_lm_output(step_model_table), file = "html",
      caption = "Stroke Model Using Stepwise Covariates and Robust Variance Estimator")
step_final_table
```

Final model suggested is: 
$$\text{Length of admission days} = \beta_0 + \beta_1\text{AdmissionNIHSS} + \beta_2\text{AdmissionMRS} + \beta_3\text{Race}+ \beta_4\text{Cholesterol}+ \beta_5\text{Age/10yrs} + \epsilon$$

Will potentially use a robust variance estimator. First need to see what residual plot looks like after multiple imputation via MICE package. 

# Interaction terms

```{r Full model with interactions}
mdl1_interactions <-
  lm(length_of_admission_days ~ admission_nihss + admission_mrs_score + 
       location_of_stroke + caregiver + prior_stroke + race + sex_male + 
       bmi_kg_m_2 + cholesterol_mg_d_l + dm + antidep_arrival + anxio_arrival +
       I((age - mean(age))/10) + statin_arrival + htn + ischemic_heart_disease +
       admission_nihss:admission_mrs_score + bmi_kg_m_2:cholesterol_mg_d_l +
       admission_nihss:location_of_stroke, 
     data = mdl1_data_complete)
summary(mdl1_interactions)
plot(mdl1_interactions)
```

# Ignoring meds at arrival

$$\text{Length of admission days} = \beta_0 + \beta_1\text{AdmissionNIHSS}_i + \beta_2\text{AdmissionMRS}_i + \beta_3\text{StrokeLocation-MultipleLocations}_i + \\ \beta_4\text{StrokeLocation-Subcortical}_i +  \beta_5\text{StrokeLocation-Cerebellum}_i + \beta_6\text{StrokeLocation-Brainstem}_i + \\ \beta_7\text{CaregiverPresence}_i + \beta_8\text{PriorStroke}_i + \beta_9\text{Race-Black}_i + \beta_{10}\text{Race-Other}_i + \beta_{11}\text{Age}_i + \\ \beta_{12}\text{Sex}_i + \beta_{13}\text{BMI}_i + \beta_{14}\text{Cholesterol}_i + \beta_{15}\text{Diabetes} + \beta_{16}\text{Antidepressants}_i + \\ \beta_{17}\text{Anxiolytics}_i+  \beta_{18}\text{Hypertension} + \beta_{19}\text{IschemicHeartDisease} + \epsilon_i$$


```{r Subsetting ignore statin}
mdl1_data <- 
  stroke |> 
  mutate(
    caregiver = case_when(
      caregiver == "Present" ~ 1,
      caregiver == "Absent" ~ 0),
    prior_stroke = case_when(
      prior_stroke == "No" ~ 0,
      prior_stroke == "Yes" ~ 1),
    sex_male = ifelse(sex == "Male", 1, 0),
    dm = ifelse(dm == "No", 0, 1),
    id_number = as.numeric(id_number),
    location_of_stroke = 
      factor(location_of_stroke,
             levels = c("Cortical","Multiple Locations",
                        "Subcortical","Cerebellum",
                        "Brainstem"))) |> 
  select(id_number,length_of_admission_days, admission_nihss, admission_mrs_score, 
         location_of_stroke, caregiver, prior_stroke, race, age, sex_male, 
         bmi_kg_m_2, cholesterol_mg_d_l, dm, antidep_arrival, anxio_arrival,
         htn, ischemic_heart_disease)
mdl1_data_complete = mdl1_data[complete.cases(mdl1_data),]
```

```{r Full model excluding statin}
mdl1 <-
  lm(length_of_admission_days ~ admission_nihss + admission_mrs_score + 
       location_of_stroke + caregiver + prior_stroke + race + sex_male + 
       bmi_kg_m_2 + cholesterol_mg_d_l + dm + antidep_arrival + anxio_arrival +
       I((age - mean(age))/10) + htn + ischemic_heart_disease, 
     data = mdl1_data_complete)
summary(mdl1)
plot(mdl1)
car::ncvTest(mdl1)
```


```{r Robust Variance Estimator for Full model ignoring statins}
#Run Dan211Functions First
library(kableExtra)
full_model_table <- lmCI(mdl1, robust = TRUE)
rownames(full_model_table) <- 
  c("Intercept", "Admission NIHSS", "Admission MRS", 
    "Diff b/w Multiple locations & Cortical", 
    "Diff b/w Subcortical & Cortical",
    "Diff b/w Cerebellum & Cortical",
    "Diff b/w Brainstem & Cortical",
    "Caregiver", "Prior Stroke", "Diff b/w Black & White", 
    "Diff b/w Other & White", "Diff b/w Male & Female",
    "BMI", "Cholesterol", "Diabetes", 
    "Antidepressants at Arrival", "Anxiolytic at Arrival", 
    "Centered Age per 10 years", "Hypertension", 
    "Ischemic Heart Disease")
full_final_table <- kable(format_lm_output(full_model_table), file = "html",
      caption = "Stroke Model Using All Covariates and Robust Variance Estimator")
full_final_table
```

```{r Stepwise ignoring statin}
step(mdl1)
mdl1_step <- 
  lm(length_of_admission_days ~ admission_nihss + admission_mrs_score + 
       race + cholesterol_mg_d_l + I((age - mean(age))/10), 
     data = mdl1_data_complete)
summary(mdl1_step)
plot(mdl1_step)
```

```{r Robust Variance Estimator for Stepwise excluding statin}
#Run Dan211Functions First
library(kableExtra)
step_model_table <- lmCI(mdl1_step, robust = TRUE)
rownames(step_model_table) <- 
  c("Intercept", "Admission NIHSS", "Admission MRS", 
    "Diff b/w Black & White", "Diff b/w Other & White", 
    "Cholesterol", "Centered Age per 10 years")
step_final_table <- kable(format_lm_output(step_model_table), file = "html",
      caption = "Stroke Model Using Stepwise Covariates and Robust Variance Estimator")
step_final_table
```

Final model suggested is: 
$$\text{Length of admission days} = \beta_0 + \beta_1\text{AdmissionNIHSS}_i + \beta_2\text{AdmissionMRS}_i + \\ \beta_3\text{Race-Black}_i + \beta_4\text{Race-Other}_i + \beta_5\text{Cholesterol}_i+ \beta_6\text{Age}_i + \epsilon_i$$
# New model 4/4/25

In this model we're using the new cortical indicator variable, side of stroke, and log of admission days as the response



```{r Potential final model}
mdl1_data <- 
  stroke |> 
  mutate(
    caregiver = case_when(
      caregiver == "Present" ~ 1,
      caregiver == "Absent" ~ 0),
    prior_stroke = case_when(
      prior_stroke == "No" ~ 0,
      prior_stroke == "Yes" ~ 1),
    sex_male = ifelse(sex == "Male", 1, 0),
    dm = ifelse(dm == "No", 0, 1),
    id_number = as.numeric(id_number),
    location_of_stroke = 
      factor(location_of_stroke,
             levels = c("Cortical","Multiple Locations",
                        "Subcortical","Cerebellum",
                        "Brainstem"))) |> 
  select(id_number, admission_nihss, admission_mrs_score, 
         location_of_stroke, caregiver, prior_stroke, race, age, sex_male, 
         bmi_kg_m_2, cholesterol_mg_d_l, dm, antidep_arrival, anxio_arrival,
         htn, ischemic_heart_disease, cigarette_smoking, lesion_cortical, 
         length_of_admission_days, log_lengthhospitalstay, side_of_stroke)

mdl1_data_complete <-
  mdl1_data[complete.cases(mdl1_data),] |> 
  filter(length_of_admission_days!=0)
```
$$\ln[\text{Length of admission days}] = \beta_0 + \beta_1\text{AdmissionNIHSS}_i + \beta_2\text{AdmissionMRS}_i + \beta_3\text{Side of Stroke}_i + \\ \beta_4\text{Cortical Lesion}_i +  \beta_5\text{Caregiver Presence}_i + \beta_6\text{Prior Stroke}_i +  \beta_7\text{Race-Black}_i + \\\beta_{8}\text{Race-Other}_i + \beta_{9}\text{Age}_i + \beta_{10}\text{Sex-Male}_i + \beta_{11}\text{BMI}_i + \beta_{12}\text{Cholesterol}_i + \\ \beta_{13}\text{Diabetes} + \beta_{14}\text{Antidepressants}_i + \beta_{15}\text{Anxiolytics}_i + \beta_{16}\text{Hypertension} + \\ \beta_{17}\text{IschemicHeartDisease} + \beta_{18}\text{SmokingHistory} + \epsilon_i$$

```{r potential full model excluding statin}
mdl1 <-
  lm(log_lengthhospitalstay ~ admission_nihss + admission_mrs_score + side_of_stroke + 
       lesion_cortical + caregiver + prior_stroke + race + sex_male + 
       bmi_kg_m_2 + cholesterol_mg_d_l + dm + antidep_arrival + anxio_arrival +
       I((age - mean(age))/10) + htn + ischemic_heart_disease + cigarette_smoking #+ prior_stroke:I((age - mean(age))/10) + htn:prior_stroke + ischemic_heart_disease:prior_stroke
     , 
     data = mdl1_data_complete)
summary(mdl1)
plot(mdl1)
```

```{r Stepwise ignoring statin}
step(mdl1)
mdl1_step <- 
  lm(log_lengthhospitalstay ~ admission_nihss + admission_mrs_score + 
       cholesterol_mg_d_l + anxio_arrival + I((age - mean(age))/10) + htn, 
     data = mdl1_data_complete)
summary(mdl1_step)
plot(mdl1_step)
```