---
title: "Methodology"
author: "Lidio Jaimes Jr"
date: "2025-02-21"
output: pdf_document
---
```{r,include=FALSE}
library(tidyverse)
library(forcats)
library(table1)
```
# Research Question

A predictive model in acute stroke patient's care incorporating pre-existing conditions (diabetes, hypertension, CAD, ischemic heart disease), lifestyle factors (smoking, obesity, hypercholesterolemia), socioeconomic, demographic factors, presence of caregiver and medication use (anticoagulants, statins, antidepressants, anxiolytics) will estimate patients' length of hospital stay at initial assessment.

$$\text{Length of admission days} = \beta_0 + \beta_1\text{AdmissionNIHSS}_i + \beta_2\text{AdmissionMRS}_i + \beta_3\text{StrokeLocation-MultipleLocations}_i + \\ \beta_4\text{StrokeLocation-Subcortical}_i +  \beta_5\text{StrokeLocation-Cerebellum}_i + \beta_6\text{StrokeLocation-Brainstem}_i + \\ \beta_7\text{CaregiverPresence}_i + \beta_8\text{PriorStroke}_i + \beta_9\text{Race-Black}_i + \beta_{10}\text{Race-Other}_i + \beta_{11}\text{Age}_i + \\ \beta_{12}\text{Sex}_i + \beta_{13}\text{BMI}_i + \beta_{14}\text{Cholesterol}_i + \beta_{15}\text{Diabetes} + \beta_{16}\text{Statins}_i + \beta_{17}\text{Antidepressants}_i + \\ \beta_{18}\text{Anxiolytics}_i + \beta_{19}\text{Hypertension} + \beta_{20}\text{IschemicHeartDisease} + \beta_{21}\text{SmokingHistory} + \epsilon_i$$

Shorter version:
$$E[\text{length of admission days}] = \beta_0 + \beta_1\text{AdmissionNIHSS}_i + \beta_2\text{AdmissionMRS}_i + \overrightarrow{\boldsymbol{\gamma}}^T\overrightarrow{\boldsymbol{z}_i}$$

What is the variable for CAD?

# Model 1 after second year of data entry (want to see meds at arrival)

```{r Subsetting complete}
mdl1_data <- 
  stroke |> 
  mutate(
    caregiver = case_when(
      caregiver == "Present" ~ 1,
      caregiver == "Absent" ~ 0),
    prior_stroke = case_when(
      prior_stroke == "No" ~ 0,
      prior_stroke == "Yes" ~ 1),
    sex_male = ifelse(sex == "Male", 1, 0),
    dm = ifelse(dm == "No", 0, 1),
    id_number = as.numeric(id_number),
    location_of_stroke = 
      factor(location_of_stroke,
             levels = c("Cortical","Multiple Locations",
                        "Subcortical","Cerebellum",
                        "Brainstem"))) |>
  filter(id_number>=375) |> #, #want everyone after second year of entry
  select(id_number,length_of_admission_days, admission_nihss, admission_mrs_score, 
         location_of_stroke, caregiver, prior_stroke, race, age, sex_male, 
         bmi_kg_m_2, cholesterol_mg_d_l, dm, antidep_arrival, anxio_arrival,
         statin_arrival, htn, ischemic_heart_disease)
```

```{r Full model base}
mdl1 <-
  lm(length_of_admission_days ~ admission_nihss + admission_mrs_score + 
       location_of_stroke + caregiver + prior_stroke + race + sex_male + 
       bmi_kg_m_2 + cholesterol_mg_d_l + dm + antidep_arrival + anxio_arrival +
       I((age - mean(age))/10) + statin_arrival + htn + ischemic_heart_disease, 
     data = mdl1_data_complete)
summary(mdl1)
plot(mdl1)
```

```{r Robust Variance Estimator for Full model}
#Run Dan211Functions First
library(kableExtra)
full_model_table <- lmCI(mdl1, robust = TRUE)
rownames(full_model_table) <- 
  c("Intercept", "Admission NIHSS", "Admission MRS", 
    "Diff b/w Multiple locations & Cortical", 
    "Diff b/w Subcortical & Cortical",
    "Diff b/w Cerebellum & Cortical",
    "Diff b/w Brainstem & Cortical",
    "Caregiver", "Prior Stroke", "Diff b/w Black & White", 
    "Diff b/w Other & White", "Diff b/w Male & Female",
    "BMI", "Cholesterol", "Diabetes", 
    "Antidepressants at Arrival", "Anxiolytic at Arrival", 
    "Centered Age per 10 years", "Statin at Arrival", "Hypertension", 
    "Ischemic Heart Disease")
full_final_table <- kable(format_lm_output(full_model_table), file = "html",
      caption = "Stroke Model Using All Covariates and Robust Variance Estimator")
full_final_table
```

```{r Stepwise}
#step(mdl1)
mdl1_step <- 
  lm(length_of_admission_days ~ admission_nihss + admission_mrs_score + 
       race + cholesterol_mg_d_l + I((age - mean(age))/10), 
     data = mdl1_data_complete)
summary(mdl1_step)
plot(mdl1_step)
```

```{r Robust Variance Estimator for Stepwise}
#Run Dan211Functions First
library(kableExtra)
step_model_table <- lmCI(mdl1_step, robust = TRUE)
rownames(step_model_table) <- 
  c("Intercept", "Admission NIHSS", "Admission MRS", 
    "Diff b/w Black & White", "Diff b/w Other & White", 
    "Cholesterol", "Centered Age per 10 years")
step_final_table <- kable(format_lm_output(step_model_table), file = "html",
      caption = "Stroke Model Using Stepwise Covariates and Robust Variance Estimator")
step_final_table
```

Final model suggested is: 
$$\text{Length of admission days} = \beta_0 + \beta_1\text{AdmissionNIHSS} + \beta_2\text{AdmissionMRS} + \beta_3\text{Race}+ \beta_4\text{Cholesterol}+ \beta_5\text{Age/10yrs} + \epsilon$$

Will potentially use a robust variance estimator. First need to see what residual plot looks like after multiple imputation via MICE package. 

# Interaction terms

```{r Full model with interactions}
mdl1_interactions <-
  lm(length_of_admission_days ~ admission_nihss + admission_mrs_score + 
       location_of_stroke + caregiver + prior_stroke + race + sex_male + 
       bmi_kg_m_2 + cholesterol_mg_d_l + dm + antidep_arrival + anxio_arrival +
       I((age - mean(age))/10) + statin_arrival + htn + ischemic_heart_disease +
       admission_nihss:admission_mrs_score + bmi_kg_m_2:cholesterol_mg_d_l +
       admission_nihss:location_of_stroke, 
     data = mdl1_data_complete)
summary(mdl1_interactions)
plot(mdl1_interactions)
```

# Ignoring meds at arrival

$$\text{Length of admission days} = \beta_0 + \beta_1\text{AdmissionNIHSS}_i + \beta_2\text{AdmissionMRS}_i + \beta_3\text{StrokeLocation-MultipleLocations}_i + \\ \beta_4\text{StrokeLocation-Subcortical}_i +  \beta_5\text{StrokeLocation-Cerebellum}_i + \beta_6\text{StrokeLocation-Brainstem}_i + \\ \beta_7\text{CaregiverPresence}_i + \beta_8\text{PriorStroke}_i + \beta_9\text{Race-Black}_i + \beta_{10}\text{Race-Other}_i + \beta_{11}\text{Age}_i + \\ \beta_{12}\text{Sex}_i + \beta_{13}\text{BMI}_i + \beta_{14}\text{Cholesterol}_i + \beta_{15}\text{Diabetes} + \beta_{16}\text{Antidepressants}_i + \\ \beta_{17}\text{Anxiolytics}_i+  \beta_{18}\text{Hypertension} + \beta_{19}\text{IschemicHeartDisease} + \epsilon_i$$


```{r Subsetting ignore statin}
mdl1_data <- 
  stroke |> 
  mutate(
    caregiver = case_when(
      caregiver == "Present" ~ 1,
      caregiver == "Absent" ~ 0),
    prior_stroke = case_when(
      prior_stroke == "No" ~ 0,
      prior_stroke == "Yes" ~ 1),
    sex_male = ifelse(sex == "Male", 1, 0),
    dm = ifelse(dm == "No", 0, 1),
    id_number = as.numeric(id_number),
    location_of_stroke = 
      factor(location_of_stroke,
             levels = c("Cortical","Multiple Locations",
                        "Subcortical","Cerebellum",
                        "Brainstem"))) |> 
  select(id_number,length_of_admission_days, admission_nihss, admission_mrs_score, 
         location_of_stroke, caregiver, prior_stroke, race, age, sex_male, 
         bmi_kg_m_2, cholesterol_mg_d_l, dm, antidep_arrival, anxio_arrival,
         htn, ischemic_heart_disease)
mdl1_data_complete = mdl1_data[complete.cases(mdl1_data),]
```

```{r Full model excluding statin}
mdl1 <-
  lm(length_of_admission_days ~ admission_nihss + admission_mrs_score + 
       location_of_stroke + caregiver + prior_stroke + race + sex_male + 
       bmi_kg_m_2 + cholesterol_mg_d_l + dm + antidep_arrival + anxio_arrival +
       I((age - mean(age))/10) + htn + ischemic_heart_disease, 
     data = mdl1_data_complete)
summary(mdl1)
plot(mdl1)
car::ncvTest(mdl1)
```


```{r Robust Variance Estimator for Full model ignoring statins}
#Run Dan211Functions First
library(kableExtra)
full_model_table <- lmCI(mdl1, robust = TRUE)
rownames(full_model_table) <- 
  c("Intercept", "Admission NIHSS", "Admission MRS", 
    "Diff b/w Multiple locations & Cortical", 
    "Diff b/w Subcortical & Cortical",
    "Diff b/w Cerebellum & Cortical",
    "Diff b/w Brainstem & Cortical",
    "Caregiver", "Prior Stroke", "Diff b/w Black & White", 
    "Diff b/w Other & White", "Diff b/w Male & Female",
    "BMI", "Cholesterol", "Diabetes", 
    "Antidepressants at Arrival", "Anxiolytic at Arrival", 
    "Centered Age per 10 years", "Hypertension", 
    "Ischemic Heart Disease")
full_final_table <- kable(format_lm_output(full_model_table), file = "html",
      caption = "Stroke Model Using All Covariates and Robust Variance Estimator")
full_final_table
```

```{r Stepwise ignoring statin}
step(mdl1)
mdl1_step <- 
  lm(length_of_admission_days ~ admission_nihss + admission_mrs_score + 
       race + cholesterol_mg_d_l + I((age - mean(age))/10), 
     data = mdl1_data_complete)
summary(mdl1_step)
plot(mdl1_step)
```

```{r Robust Variance Estimator for Stepwise excluding statin}
#Run Dan211Functions First
library(kableExtra)
step_model_table <- lmCI(mdl1_step, robust = TRUE)
rownames(step_model_table) <- 
  c("Intercept", "Admission NIHSS", "Admission MRS", 
    "Diff b/w Black & White", "Diff b/w Other & White", 
    "Cholesterol", "Centered Age per 10 years")
step_final_table <- kable(format_lm_output(step_model_table), file = "html",
      caption = "Stroke Model Using Stepwise Covariates and Robust Variance Estimator")
step_final_table
```

Final model suggested is: 
$$\text{Length of admission days} = \beta_0 + \beta_1\text{AdmissionNIHSS}_i + \beta_2\text{AdmissionMRS}_i + \\ \beta_3\text{Race-Black}_i + \beta_4\text{Race-Other}_i + \beta_5\text{Cholesterol}_i+ \beta_6\text{Age}_i + \epsilon_i$$
# New model 4/4/25

In this model we're using the new cortical indicator variable, side of stroke, and log of admission days as the response



```{r Potential final model}
mdl1_data <- 
  stroke |> 
  mutate(
    caregiver = case_when(
      caregiver == "Present" ~ 1,
      caregiver == "Absent" ~ 0),
    prior_stroke = case_when(
      prior_stroke == "No" ~ 0,
      prior_stroke == "Yes" ~ 1),
    sex_male = ifelse(sex == "Male", 1, 0),
    dm = ifelse(dm == "No", 0, 1),
    id_number = as.numeric(id_number),
    location_of_stroke = 
      factor(location_of_stroke,
             levels = c("Cortical","Multiple Locations",
                        "Subcortical","Cerebellum",
                        "Brainstem"))) |> 
  select(id_number, admission_nihss, admission_mrs_score, 
         location_of_stroke, caregiver, prior_stroke, race, age, sex_male, 
         bmi_kg_m_2, cholesterol_mg_d_l, dm, antidep_arrival, anxio_arrival,
         htn, ischemic_heart_disease, cigarette_smoking, lesion_cortical, 
         length_of_admission_days, log_lengthhospitalstay, side_of_stroke)

mdl1_data_complete <-
  mdl1_data[complete.cases(mdl1_data),] |> 
  filter(length_of_admission_days!=0)
```
$$\ln[\text{Length of admission days}] = \beta_0 + \beta_1\text{AdmissionNIHSS}_i + \beta_2\text{AdmissionMRS}_i + \beta_3\text{Side of Stroke}_i + \\ \beta_4\text{Cortical Lesion}_i +  \beta_5\text{Caregiver Presence}_i + \beta_6\text{Prior Stroke}_i +  \beta_7\text{Race-Black}_i + \\\beta_{8}\text{Race-Other}_i + \beta_{9}\text{Age}_i + \beta_{10}\text{Sex-Male}_i + \beta_{11}\text{BMI}_i + \beta_{12}\text{Cholesterol}_i + \\ \beta_{13}\text{Diabetes} + \beta_{14}\text{Antidepressants}_i + \beta_{15}\text{Anxiolytics}_i + \beta_{16}\text{Hypertension} + \\ \beta_{17}\text{IschemicHeartDisease} + \beta_{18}\text{SmokingHistory} + \epsilon_i$$

```{r potential full model excluding statin}
mdl1 <-
  lm(log_lengthhospitalstay ~ admission_nihss + admission_mrs_score + side_of_stroke + 
       lesion_cortical + caregiver + prior_stroke + race + sex_male + 
       bmi_kg_m_2 + cholesterol_mg_d_l + dm + antidep_arrival + anxio_arrival +
       I((age - mean(age))/10) + htn + ischemic_heart_disease + cigarette_smoking #+ prior_stroke:I((age - mean(age))/10) + htn:prior_stroke + ischemic_heart_disease:prior_stroke
     , 
     data = mdl1_data_complete)
summary(mdl1)
plot(mdl1)
```

```{r Stepwise ignoring statin}
step(mdl1)
mdl1_step <- 
  lm(log_lengthhospitalstay ~ admission_nihss + admission_mrs_score + 
       cholesterol_mg_d_l + anxio_arrival + I((age - mean(age))/10) + htn, 
     data = mdl1_data_complete)
summary(mdl1_step)
plot(mdl1_step)
```